\documentclass{customdocclass}

\title{ASCII-Serial-Com Data Format}
\author{Justin Hugon}
\date{2021-11-19}

\begin{document}

\maketitle

\section{Version 0}

\tredbf{Need to be careful of the bit-width of register numbers!}

This is the format for a serial communication protocol based on ASCII. The
protocol is not meant to be efficient, but to be easily human readable.

\textbf{This protocol defines a host and a device (like USB). Command types
have different meanings depending on if they are sent from a host or device}

\subsection{Data Frame}

Data frames are made up of 8-bit bytes. Only the following symbols are present in frames:

\vspace{1em}

a-zA-Z0-9,.> newline (\textbackslash n) and space

\vspace{1em}

All of the characters are 8-bit ASCII (with the MSB set to 0).

The maximum frame size is 64 bytes to reduce buffer size in microcontrollers.

\subsubsection{Data Frame Format}

The data frame consists of the following:

\begin{itemize}
  \item The start of frame character `>'
  \item One byte of version information for ASCII Serial Com; must be ASCII 0
  \item One byte of application version and format information represented in the characters A-Z0-9
  \item One byte of command and/or message type as lower case letters a-z
  \item Data is then sent as hex 0-9A-F (the letters must be capital)
  \begin{itemize}
    \item The protocol may use the comma character `,' to separate fields
    \item Applications may use the space character to group data
    \item 54 or less bytes
  \end{itemize}
  \item The end of data character `.'
  \item A CRC code represented as 4 bytes of ASCII hex 0-9A-F (capitals), so from 0000-FFFF. The CRC should be run from the start of frame character to the end of data character (inclusive). Use CRC-16-DNP (\url{http://reveng.sourceforge.net/crc-catalogue/16.htm\#crc.cat-bits.16}).
  \item The end of frame character `\textbackslash n'
\end{itemize}

Examples:

\begin{lstlisting}
>00r0F.9AD2\n
>02w0F 003FFF92.EA89\n
\end{lstlisting}

\subsection{Command and/or Message Types}

\begin{center}
\small
Command/Message Type Summary Table
\begin{tabularx}{\textwidth}{|X|l|X|X|} \hline
\multicolumn{2}{|c|}{Command/Type} & & \\ \hline
Description & Code & Host Sent Data & Dev Sent Data \\ \hline
Read register & r & Reg num (4-byte) & Reg num (4-byte) comma (`,') reg content \\ \hline
Write register & w & Reg num (4-byte) comma (`,') reg content & Reg num (4-byte) for success \\ \hline
Data stream frame & s & frame number (2-byte) comma (`,') data &  frame number (2-byte) comma (`,') data \\ \hline
Device stream on & n & & \\ \hline
Device stream off & f & & \\ \hline
Error message & e & 2 hex digits of error code, comma (`,'), command code, comma (`,'), first 9 bytes of data of message that caused error & 2 hex digits of error code, comma (`,'), command code, comma (`,'), first 9 bytes of data of message that caused error \\ \hline
\end{tabularx}
\end{center}

\begin{itemize}
  \item r: Read register: The host sends this command with a register number as 4-bytes of data in capital hex (so from 0000-FFFF). The device will respond with an `r' message with data register number in 4-bytes then comma (`,') then register value, both in capital hex.
  \item w: Write register: The host sends this command with data: register number as 4-bytes of data (so from 0000-FFFF) then comma (`,') then register value, both in capital hex. On success, the device will respond with a `w' message with the register number as 4-bytes of data (in capital hex).
  \item s: Data stream frame: The same format is sent from the host to the device or the device to the host. For streaming from the host, the host just starts streaming. For streaming to the host, the host must send the `n' and `f' commands to start and stop streaming. The data consists of a stream frame number, a `,' character, and the data. The stream frame number and data must be 2 bytes of ASCII capital hex (so from 00-FF). Spaces may be put into the data to act as e.g. field separators. The stream frame number must increment by one with each frame sent.
  \item n: Device stream on: The host sends this with no data to tell the device to start streaming data.
  \item f: Device stream off: The host sends this with no data to tell the device to stop streaming data.
  \item e: Usually device to host, replying with error and beginning of message that caused error.
\end{itemize}

\begin{center}
\small
Error Code Table
\begin{tabularx}{\textwidth}{|l|X|} \hline
Code & Description \\ \hline
0x00 & Unknown Error \\ \hline
0x01 & No Error \\ \hline
0x10 & Data too long for frame \\ \hline
0x11 & Problem computing checksum \\ \hline
0x12 & Invalid frame \\ \hline
0x13 & Invalid frame: missing or misplaced `.' \\ \hline
0x14 & Invalid frame: non-hex char where hex char expected \\ \hline
0x15 & Command not implemented \\ \hline
0x16 & Unexpected command \\ \hline
0x17 & Data too short \\ \hline
0x20 & Register block is null (invalid) \\ \hline
0x21 & Register number out of bounds \\ \hline
0x22 & Register value is wrong number of bytes \\ \hline
0x30 & Circular buffer out of bounds \\ \hline
0x31 & Circular buffer tried to pop from empty buffer \\ \hline
0x40 & File read error \\ \hline
0x41 & File write error \\ \hline
\end{tabularx}
\end{center}


\end{document}
